---
title: "lab 3"
author: "Linnea Powell"
date: "February 9, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F ) 
```

# load and plot shapefile
```{r}
library( maptools )
library( sp )

syr <- readShapePoly( fn="01-05-2015", proj4string=CRS("+proj=longlat +datum=WGS84") )

plot( syr,  border="gray80" )
```

# grab census data
```{r}
devtools::install_github("hrecht/censusapi")
censuskey <- "b06146bd7469912826a96e117fcf4a8ab8bc36de"
library(censusapi)
```

# select variables Per Capita Income, Asian population, Income < poverty level last 12 months
```{r}
myvars <- c("NAME", "B19301_001E", "B01001D_001E", "B17001_002E")
acs5_2014 <- getCensus(name="acs5", vintage=2014, key=censuskey,
    vars=myvars, region="tract:*", regionin="state:36")
head( acs5_2014 )
oc <- subset(acs5_2014, county == "067")
oc$newtract <- (as.numeric(oc$tract)/100)
head(oc)

```

#Variable 1
# making color pallette for Per Capita Income
```{r}
color.function <- colorRampPalette( c("light gray","steel blue" ) )

col.ramp <- color.function( 4 )

color.vector <- cut( rank(oc$B19301_001E), breaks=4, label=col.ramp)

color.vector <- as.character( color.vector )


```

# match up shapefile and census data
```{r}
this.order <- match(syr$CensusTrac, oc$newtract )

color.vec.ordered <- color.vector[ this.order ]

```

# get feel of Per Capita income (B19301_001E)
```{r}
summary(oc$B19301_001E)
hist(oc$B19301_001E)
```

# Graph Per Capita Income in Syracuse
```{r}

plot( syr, col=color.vec.ordered, border=FALSE, main="Syracuse Per Capita Income, 2014" )

legend.text=c("less than $20,250" ,
              "$20,251 - $28,680",
              "$28,681 - $34,380",
              "more than $34,381")

legend( "bottomleft", bg="white",
        pch=20, pt.cex=1.5, cex=0.8,
        legend=legend.text, 
        col=col.ramp, 
        box.col="white",
        title="Per Capita Income" 
       )
      
```

#Variable 2
# making color pallette for Asian population
```{r}
color.function <- colorRampPalette( c("light gray","steel blue" ) )

col.ramp <- color.function( 4 )

color.vector2 <- cut( rank(oc$B01001D_001E), breaks=4, label=col.ramp)

color.vector2 <- as.character( color.vector2 )

this.order <- match(syr$CensusTrac, oc$newtract )

color.vec.ordered2 <- color.vector2[ this.order ]

```


# get feel of Asian population (B01001D_001E)
```{r}
summary(oc$B01001D_001E)
hist(oc$B01001D_001E)
```

# Graph Asian population in Syracuse
```{r}

plot( syr, col=color.vec.ordered2, border=FALSE, main="Asian population, 2014" )

legend.text=c("less than 10" ,
              "10-46",
              "47-143",
              "144+")

legend( "bottomleft", bg="white",
        pch=20, pt.cex=1.5, cex=0.8,
        legend=legend.text, 
        col=col.ramp, 
        box.col="white",
        title="Number of Asian people" 
       )
      
```


#Variable 3
# making color pallette for Income < poverty level last 12 months
```{r}
color.function <- colorRampPalette( c("light gray","red" ) )

col.ramp <- color.function( 4 )

color.vector3 <- cut( rank(oc$B17001_002E), breaks=4, label=col.ramp)


color.vector3 <- as.character( color.vector3 )

this.order <- match(syr$CensusTrac, oc$newtract )

color.vec.ordered3 <- color.vector3[ this.order ]

```


# get feel of Income < poverty level last 12 months (B17001_002E)
```{r}
summary(oc$B17001_002E)
hist(oc$B17001_002E)
```

# Graph Income < poverty level last 12 months in Syracuse
```{r}

plot( syr, col=color.vec.ordered3, border=FALSE, main="Income below poverty level in Syracuse, 2014" )

legend.text=c("0-180" ,
              "180-343",
              "343-674",
              "675+")

legend( "bottomright", bg="white",
        pch=20, pt.cex=1.5, cex=0.8,
        legend=legend.text, 
        col=col.ramp, 
        box.col="white",
        title="Number of people with \n income below poverty level" 
       )
      
```