---
title: "Lab 6"
author: "Linnea Powell"
date: "March 9, 2017"
output: github_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set( message=F, warning=F )
```

### load libraries
```{r}
library( maptools )
library( sp )
library(raster)
library(rgeos)
library(dplyr)
library( spatialEco )
library( rgdal )
library( geojsonio )
library(maps)
```

## Part 1

### download onondaga census tracts
```{r}
download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2010/TRACT/2010/tl_2010_36067_tract10.zip", "onondaga census tracts.zip" )

unzip( "onondaga census tracts.zip" )

file.remove( "onondaga census tracts.zip" )

#dir()
```

### load and plot your shapefile
```{r}
syr <- readShapePoly( fn="tl_2010_36067_tract10", proj4string=CRS("+proj=longlat +datum=WGS84") )

par( mar=c(0,0,0,0) )
plot( syr,  border="gray10" )

```

### add roads
```{r}
download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2015/PRISECROADS/tl_2015_36_prisecroads.zip", "primary_and_secondary_roads.zip" )

unzip( "primary_and_secondary_roads.zip" )

file.remove( "primary_and_secondary_roads.zip" )

roads <- readShapeLines( fn="tl_2015_36_prisecroads", proj4string=CRS("+proj=longlat +datum=WGS84") )

head( roads@data )
table( roads$RTTYP )
```

### extract interstates
```{r}
par( mar=c(0,0,0,0) )
plot( syr,  border="gray80" )
plot( roads, col="steelblue", lwd=2, add=T )

interstate <- roads[ roads$RTTYP == "I" , ]
plot( interstate, col="red", add=T )
```

### clip shapefile
```{r}
gClip <- function(shp, bb){
if(class(bb) == "matrix") b_poly <- as(extent(as.vector(t(bb))), "SpatialPolygons")
else b_poly <- as(extent(bb), "SpatialPolygons")
gIntersection(shp, b_poly, byid = T)
}

roads_clipped <- gClip(roads, syr)

par( mar=c(0,0,0,0) )
plot( syr,  border="gray80" )
plot( roads_clipped, col="steelblue", lwd=2, add=T )
```

### extract interstates as a separate layer 
```{r}
interstate <- roads[ roads$RTTYP == "I" , ]
```

### create interstate buffer
```{r}
interstate_clipped <- gClip(interstate, syr)

par( mar=c(0,0,0,0) )
plot( syr, border="gray80" )
plot( interstate_clipped, col="red", add=T)

buff1 <- gBuffer(interstate_clipped, width = .005, byid=FALSE)
plot(buff1, add=T)

```

### identify all houses within zone and add new variable
```{r}
house.dat <- read.csv("https://raw.githubusercontent.com/lecy/hedonic-prices/master/Data/Housing%20Price%20In-Class%20Exercise%20(Responses).csv")
#head(house.dat)
lat.long <- read.csv ("https://raw.githubusercontent.com/lecy/hedonic-prices/master/Data/lat.long.csv")

lat.long2<-SpatialPoints(lat.long, proj4string=CRS("+proj=longlat +datum=WGS84"))
matchpointstobuffer <- over(lat.long2, buff1 )
matchpointstobuffer[is.na(matchpointstobuffer)] <- 0
inZone <-matchpointstobuffer==1
inZone
dathouse <- as.data.frame(house.dat )
#dat.buffer <- cbind( dathouse, inZone)#problem with differing number of rows

house.dat.buffer<- merge(dathouse, inZone)

plot( syr, border="gray80" )
plot(buff1, col="red", add=T)
points(lat.long2$lon[inZone], lat.long2$lat[inZone], col="blue", pch=19, cex=.8)

```

##Part 2

### load geojson file
```{r}
syr.geo <- readOGR(dsn="C:/Users/powel/Documents/Data 2/syr_parcels.geojson", layer="OGRGeoJSON") 

```

### Industrial Zones - create buffer, plot it, identify houses within it, create new variable

```{r}
industrial_zone <- syr.geo[syr.geo$LandUse=="Industrial", ]
other_zones <- syr.geo[syr.geo$LandUse!="Industrial", ]

plot(other_zones, col = "grey", border = F)
plot(industrial_zone, add = T, col = "blue", border = F)

industrialBuffer <- gBuffer(industrial_zone, width = .005)
plot(industrialBuffer, add = T, col = "yellow")
points(lat.long2$lon, lat.long2$lat, col="blue", pch=19, cex=.5)

industrialBuffer.proj<- spTransform(industrialBuffer, CRS("+proj=longlat +datum=WGS84"))
industrial_over <- over(lat.long2, industrialBuffer.proj)

industrial_over[is.na(industrial_over)] <- 0
in_industrial_zone <-industrial_over==1
dathouse_with_indzone <- merge(dathouse, in_industrial_zone)
head(dathouse_with_indzone)
names(dathouse_with_indzone)[names(dathouse_with_indzone)=="y"] <- "In Industrial Zone"
head(dathouse_with_indzone)


```



### Schools - create buffer, plot it, identify houses within it, create new variable
```{r}
schools <- syr.geo[syr.geo$LandUse=="Schools", ]
not_schools <- syr.geo[syr.geo$LandUse!="Schools", ]

plot(not_schools, col = "grey", border = F)
plot(schools, add = T, col = "blue", border = F)

schoolBuffer <- gBuffer(schools, width = .0025)
plot(schoolBuffer, add = T, col = "green")
points(lat.long2$lon, lat.long2$lat, col="blue", pch=19, cex=.8)

schoolBuffer.proj<- spTransform(schoolBuffer, CRS("+proj=longlat +datum=WGS84"))
schools_over <- over(lat.long2, schoolBuffer.proj)
schools_over[is.na(schools_over)] <- 0
near_school <-schools_over==1
dathouse_with_near_school <- merge(dathouse, near_school)
head(dathouse_with_near_school)
names(dathouse_with_near_school)[names(dathouse_with_near_school)=="y"] <- "Near School"
head(dathouse_with_indzone)

```

